---
title: "Lake George Forecast"
date: "`r Sys.Date()`"
output:  
  bookdown::pdf_document2:
    toc: false
    number_sections: false
    includes:  
      in_header: my_header.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
	message = FALSE,
	warning = FALSE,
	comment = NA
)
```


```{r}
library(Rnoaa4cast)
#library(LakeEnsemblR)
library(gotmtools)
library(rLakeAnalyzer)

library(ncdf4)

library(lubridate)
library(dplyr)
library(tidyr)

library(ggplot2)
library(patchwork)
```

```{r}
initial_fc_date <- Sys.Date()
```

```{r}
qfun <- function(x, p = c(0.25, 0.75)){
  return(data.frame(ymin = quantile(x, probs = p[1]), 
                    y = median(x), 
                    ymax = quantile(x, probs = p[2])))
}
```


# Meteorological 

## Wind

```{r}
metpath <- paste0(here::here(), "/drivers/met_archive/meteo_fc_", 
                  initial_fc_date, ".csv")
ens_meteo <- data.table::fread(metpath)
```


```{r, fig.cap = "Wind speed at 10m, light gray shaded area is the middle 80% of predicted values, while dark shaded area is middle 50%. Red line indicates high wind (6 m/s) and blue line indicates low wind (3 m /s)", fig.height = 4}
ens_meteo %>% 
  gather(var, value, Air_Temperature_celsius:Ten_Meter_Elevation_Wind_Speed_meterPerSecond,
         Snowfall_millimeterPerHour) %>% 
  filter(var == "Ten_Meter_Elevation_Wind_Speed_meterPerSecond") %>% 
  ggplot(aes(x = datetime, y = value)) + 
  # geom_line(aes(group = ens), size = 0.7, alpha = 0.7) + 
  # stat_summary(fun.data = mean_se,geom = "ribbon", fill = "green3") + 
  stat_summary(fun.data = qfun, fun.args = list(p = c(0.1, 0.9)), geom = "ribbon", fill = "gray70") +
  stat_summary(fun.data = qfun, geom = "ribbon", fill = "gray40") + 
  stat_summary(fun.data = qfun, geom = "line", size = 1.2) + 
  geom_hline(yintercept = 6, color = "red", size = 1.2) + 
  geom_hline(yintercept = 3, color = "blue", size = 1.2) +
  scale_x_datetime(breaks = seq(ymd_hms(paste(initial_fc_date, "00:00:00")),
                              ymd_hms(paste(initial_fc_date, "00:00:00"))+days(16), "2 days"), 
                   date_labels = "%m/%d") + 
  labs(x = "Datetime (hours)", y = "Wind Speed (m/s)") + 
  theme_bw(base_size = 14)
```



## Air Temperature  

```{r, fig.cap = "Air temperature, light gray shaded area is the middle 80% of predicted values, while dark shaded area is middle 50%.", fig.height = 4}
ens_meteo %>% 
  gather(var, value, Air_Temperature_celsius:Ten_Meter_Elevation_Wind_Speed_meterPerSecond,
         Snowfall_millimeterPerHour) %>% 
  filter(var == "Air_Temperature_celsius") %>% 
  ggplot(aes(x = datetime, y = value * 0.45)) + 
  # geom_line(aes(group = ens), size = 0.7, alpha = 0.7) + 
  # stat_summary(fun.data = mean_se,geom = "ribbon", fill = "green3") + 
  stat_summary(fun.data = qfun, fun.args = list(p = c(0.1, 0.9)), geom = "ribbon", fill = "gray70") +
  stat_summary(fun.data = qfun, geom = "ribbon", fill = "gray40") + 
  stat_summary(fun.data = qfun, geom = "line", size = 1.2) + 
  scale_x_datetime(breaks = seq(ymd_hms(paste(initial_fc_date, "00:00:00")),
                              ymd_hms(paste(initial_fc_date, "00:00:00"))+days(16), "2 days"), 
                   date_labels = "%m/%d") + 
  labs(x = "Datetime (hours)", y = "Air Temperature (°C)") + 
  theme_bw(base_size = 14)
```

## Light  

```{r, fig.cap = "Photosynthetically active radiation, defined as 0.45 times downwelling shortwave radiation, light gray shaded area is the middle 80% of predicted values, while dark shaded area is middle 50%.", fig.height = 4}
ens_meteo %>% 
  gather(var, value, Air_Temperature_celsius:Ten_Meter_Elevation_Wind_Speed_meterPerSecond,
         Snowfall_millimeterPerHour) %>% 
  filter(var == "Shortwave_Radiation_Downwelling_wattPerMeterSquared") %>% 
  ggplot(aes(x = datetime, y = value * 0.45)) + 
  # geom_line(aes(group = ens), size = 0.7, alpha = 0.7) + 
  # stat_summary(fun.data = mean_se,geom = "ribbon", fill = "green3") + 
  stat_summary(fun.data = qfun, fun.args = list(p = c(0.1, 0.9)), geom = "ribbon", fill = "gray70") +
  stat_summary(fun.data = qfun, geom = "ribbon", fill = "gray40") + 
  stat_summary(fun.data = qfun, geom = "line", size = 1.2) + 
  scale_x_datetime(breaks = seq(ymd_hms(paste(initial_fc_date, "00:00:00")),
                              ymd_hms(paste(initial_fc_date, "00:00:00"))+days(16), "2 days"), 
                   date_labels = "%m/%d") + 
  labs(x = "Datetime (hours)", y = "PAR (Watts/m2)") + 
  theme_bw(base_size = 14)
```

## Precipitation  

```{r, fig.cap = "Precipitation, light gray shaded area is the middle 80% of predicted values, while dark shaded area is middle 50%.", fig.height = 4}
ens_meteo %>% 
  gather(var, value, Air_Temperature_celsius:Ten_Meter_Elevation_Wind_Speed_meterPerSecond,
         Snowfall_millimeterPerHour) %>% 
  filter(var == "Rainfall_millimeterPerHour") %>% 
  ggplot(aes(x = datetime, y = value * 0.45)) + 
  # geom_line(aes(group = ens), size = 0.7, alpha = 0.7) + 
  # stat_summary(fun.data = mean_se,geom = "ribbon", fill = "green3") + 
  stat_summary(fun.data = qfun, fun.args = list(p = c(0.1, 0.9)), geom = "ribbon", fill = "gray70") +
  stat_summary(fun.data = qfun, geom = "ribbon", fill = "gray40") + 
  stat_summary(fun.data = qfun, geom = "line", size = 1.2) + 
  scale_x_datetime(breaks = seq(ymd_hms(paste(initial_fc_date, "00:00:00")),
                              ymd_hms(paste(initial_fc_date, "00:00:00"))+days(16), "2 days"), 
                   date_labels = "%m/%d") + 
  labs(x = "Datetime (hours)", y = "Precipitation (mm/hr)") + 
  theme_bw(base_size = 14)
```


# Lake George water  

## 1D Temperature  

```{r}
ltpath <- paste0(here::here(), "/temp_forecasts/fc_", initial_fc_date, ".csv")
laketemps <- data.table::fread(ltpath)
```

```{r, fig.cap = "Mean estimated water temperature profiles", fig.height = 7}
laketemps %>% group_by(Var1, Var2, Var3) %>% 
  summarize(value = mean(value)) %>% 
  ggplot(aes(x = ymd_hms(Var2), y = -Var3, fill = value)) + 
  geom_raster() + #scale_fill_continuous(limits = c(0,25)) + 
  scale_y_reverse() + 
  scale_x_datetime(breaks = seq(ymd_hms(paste(initial_fc_date, "00:00:00")),
                              ymd_hms(paste(initial_fc_date, "00:00:00"))+days(16), "4 days"), 
                   date_labels = "%m/%d") + 
  facet_wrap(~Var1, ncol = 1) +
  scale_fill_viridis_c(option = "B") + 
  theme_bw(base_size = 14) + 
  theme(strip.background = element_rect(fill = "white")) +
  labs(x = "Datetime (hours)", y = "Depth (m)", fill = "°C")
```


## Thermocline 

```{r, fig.cap = "Median overall estimated thermocline depth with uncertainty"}
laketemps %>% group_by(Var1, Var2, ens) %>% 
  summarize(thermo = thermo.depth(value, -Var3, seasonal = FALSE, Smin = 2)) %>% 
  ggplot(aes(x = ymd_hms(Var2), y = thermo)) + 
  stat_summary(fun.data = qfun, geom = "ribbon", fill = "firebrick4", alpha = 0.8) +
  stat_summary(fun.data = qfun, fun.args = list(p = c(0.1, 0.9)),
               geom = "ribbon", fill = "firebrick3", alpha = 0.4) +
  stat_summary(fun = median, geom = "line", size = 1.2) + 
  #facet_wrap(~Var1, ncol = 1) + 
  scale_y_reverse() + 
  scale_x_datetime(breaks = seq(ymd_hms(paste(initial_fc_date, "00:00:00")),
                              ymd_hms(paste(initial_fc_date, "00:00:00"))+days(16), "2 days"), 
                   date_labels = "%m/%d") + 
  theme_bw(base_size = 14) + 
  labs(x = "Datetime (hours)", y = "Depth (m)")
```



```{r, fig.cap = "Estimated thermocline depth by model with uncertainty", fig.height = 8}
laketemps %>% group_by(Var1, Var2, ens) %>% 
  summarize(thermo = thermo.depth(value, -Var3, seasonal = FALSE, Smin = 2)) %>% 
  ggplot(aes(x = ymd_hms(Var2), y = thermo)) + 
  stat_summary(fun.data = qfun, geom = "ribbon", fill = "firebrick4", alpha = 0.8) +
  stat_summary(fun.data = qfun, fun.args = list(p = c(0.1, 0.9)),
               geom = "ribbon", fill = "firebrick3", alpha = 0.4) +
  #geom_point() + 
  facet_wrap(~Var1, ncol = 1) + 
  scale_y_reverse() + 
  scale_x_datetime(breaks = seq(ymd_hms(paste(initial_fc_date, "00:00:00")),
                              ymd_hms(paste(initial_fc_date, "00:00:00"))+days(16), "2 days"), 
                   date_labels = "%m/%d") + 
  theme_bw(base_size = 14) + 
  labs(x = "Datetime (hours)", y = "Depth (m)")
```

\newpage

## Comparison of previous forecast to observations

```{r, eval = FALSE}

strtdt <- Sys.Date() - 8
enddt <- Sys.Date() -1
jpvp_baseurl <- "http://jp-was.watson.ibm.com/AssemblyPoint/Data/DeepThunderAssimilationReport.csv?sid=43"
strtend <- paste0("start=", stringr::str_replace_all(strtdt, "-", ""), "000000", "&end=", stringr::str_replace_all(enddt, "-", ""), "000000")
filename <- paste0("fileName=JP_VP_006A_ALL-Variables_", strtdt, "_", enddt)

obsvp <- read.csv(paste(jpvp_baseurl, strtend, filename, sep = "&"), skip = 10, na.strings = "-")

wt <- obsvp %>% select(MM.dd.yyyy, HH.mm.ss, PC.RFU.Water.EXO..RFU., Chla.RFU.Water.EXO..RFU.,
                       Temperature.Water.EXO..C., DO.Water.EXO..mg.L., RS232SensorDepth.Water.EXO..m.) 
colnames(wt) <- c("date", "time", "pc_rfu", "chla_rfu", "temp_c", "do_mgL", "depth_m")

obswt <- wt %>% mutate(datetime = mdy_hms(paste(date, time), tz = "UTC")) %>% 
  mutate(datetime = floor_date(datetime, "1 hour"), depth_m = plyr::round_any(depth_m, 0.5)) %>% 
  ggplot(aes(x = datetime, y = depth_m, fill = temp_c)) + geom_tile() + 
  scale_fill_viridis_c(option = "B", limits = c(0, max(wt$temp_c, na.rm =TRUE))) + scale_y_reverse()

meanpred <- laketemps %>% group_by(Var2, Var3) %>% 
  summarize(value = median(value, na.rm = TRUE)) %>% 
  filter(Var2 < enddt, Var2 > strtdt) %>% 
  ggplot(aes(x = ymd_hms(Var2), y = -Var3, fill = value)) + 
  geom_raster() + #scale_fill_continuous(limits = c(0,25)) + 
  scale_y_reverse(limits = c(10.6, 0)) + 
  scale_x_datetime(breaks = seq(ymd_hms(paste(initial_fc_date, "00:00:00")),
                              ymd_hms(paste(initial_fc_date, "00:00:00"))+days(16), "4 days"), 
                   date_labels = "%m/%d") + 
  scale_fill_viridis_c(option = "B", limits = c(0, max(wt$temp_c, na.rm =TRUE))) + 
  theme_bw(base_size = 14) + 
  theme(strip.background = element_rect(fill = "white")) +
  labs(x = "Datetime (hours)", y = "Depth (m)", fill = "°C")

obswt / meanpred
```


```{r, eval = FALSE}
wt2 <- wt %>% mutate(datetime = mdy_hms(paste(date, time), tz = "UTC")) %>% 
  mutate(datetime = floor_date(datetime, "1 hour"), Var3 = -round(depth_m)) %>% 
  filter(!is.na(temp_c)) %>% 
  group_by(datetime, Var3) %>% summarize(temp_c = mean(temp_c))
  
#wt %>% mutate(Var2 = paste(date, time, sep = " "), Var3 = -round(depth_m))
laketemps %>% left_join(wt2, by = c("Var2" = "datetime", "Var3")) %>% filter(!is.na(temp_c)) %>% 
  ggplot(aes(x = temp_c, y = value)) + geom_point() + 
  geom_abline(intercept = 0, slope = 1) +# facet_wrap(~Var1) + 
  theme_bw() + 
  labs(x = "Observed Temp (°C)", y = "Mean Temp Prediction (°C)")

```


\newpage

## (BETA) Food web  

> **NOTE: this section is still in development and should not be used for decision making**  


### ATN: Allometric Trophic Network  

A box model of Lake George to determine mean daily biomass of phytoplankton and zooplankton groups

```{r}
atn_out <- read.csv(paste0(here::here(), "/ATN/output/atn_fc.csv"))
```


```{r}
atn_out %>% 
  filter(Var2 %in% c("blue", "green", "red", "brown")) %>% 
  ggplot(aes(x = Var1 + ymd("2022-04-20"), y = value, fill = Var2)) + 
  stat_summary(fun.data = qfun, fun.args = list(p = c(0.1, 0.9)), geom = "ribbon", alpha = 0.6) + 
  stat_summary(fun = median, geom = "line", aes(color = Var2)) + 
  scale_fill_manual(values = c("blue", "chocolate4", "green4", "salmon")) + 
  scale_color_manual(values = c("blue", "chocolate4", "green4", "salmon")) + 
  theme_bw() + guides(color = "none") + 
  labs(x = "Date", y = "Biomass (ug/L)", fill = "Taxa")
```


```{r}
atn_out %>% 
  filter(Var2 %in% c("rot", "clad", "cal", "cyc")) %>% 
  ggplot(aes(x = Var1 + ymd("2022-04-20"), y = value, fill = Var2)) + 
  stat_summary(fun.data = qfun, fun.args = list(p = c(0.1, 0.9)), geom = "ribbon", alpha = 0.6) + 
  stat_summary(fun = median, geom = "line", aes(color = Var2)) + 
  scale_fill_viridis_d(option = "B", end = 0.8) + 
  scale_color_viridis_d(option = "B", end = 0.8) + 
  theme_bw() + guides(color = "none") + 
  labs(x = "Date", y = "Biomass (ug/L)", fill = "Taxa")

```

\newpage

### GLM: General Lake Model  

A 1D biogeochemical model including nutrients, oxygen, and 3 phytoplankton and 1 zooplankton group

```{r}
chla <- read.csv(paste0(here::here(), "/GLMAED/archive_fc/chla.csv"))
chla %>% mutate(DateTime = ymd_hms(DateTime)) %>%  
  group_by(depth, DateTime) %>% summarize(value = median(value)) %>%
  ggplot(aes(x = DateTime, y = depth, fill = value)) + geom_raster(interpolate = TRUE) +
  scale_fill_viridis_c() + scale_y_reverse() +
  labs(title = "chla")
```

```{r}
phs <- read.csv(paste0(here::here(), "/GLMAED/archive_fc/phs.csv"))
phs %>% mutate(DateTime = ymd_hms(DateTime)) %>% 
  group_by(depth, DateTime) %>% summarize(value = median(value)) %>%
  ggplot(aes(x = DateTime, y = depth, fill = value)) + geom_raster(interpolate = TRUE) +
  scale_fill_viridis_c() + scale_y_reverse() +
  labs(title = "phosphorus")
```

```{r}
oxy <- read.csv(paste0(here::here(), "/GLMAED/archive_fc/oxy.csv"))
oxy %>% mutate(DateTime = ymd_hms(DateTime)) %>% 
  group_by(depth, DateTime) %>% summarize(value = median(value)) %>%
  ggplot(aes(x = DateTime, y = depth, fill = value)) + geom_raster(interpolate = TRUE) +
  scale_fill_viridis_c() + scale_y_reverse() +
  labs(title = "oxygen")
```



```{r}
nit <- read.csv(paste0(here::here(), "/GLMAED/archive_fc/nit.csv"))
nit %>% mutate(DateTime = ymd_hms(DateTime)) %>%  
  group_by(depth, DateTime) %>% summarize(value = median(value)) %>%
  ggplot(aes(x = DateTime, y = depth, fill = value)) + geom_raster(interpolate = TRUE) +
  scale_fill_viridis_c() + scale_y_reverse() +
  labs(title = "nitrate")
```

### WET: Water Ecosystems Tool

TBD

